---
title: "Homework 05: Logistic regression"
subtitle: "Due: Wednesday, April 21 11:59pm ET"
author: "Navya Belavadi"
date: "04-21-2021"
output:
  pdf_document: default
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, 
                      warning = FALSE, comment = "#>", highlight = TRUE)
```

# Packages

```{r packages}
library(tidyverse)
```

# Data

```{r load_data}
donner <- read_csv("data/donner.csv")
```

# Exercises

## Exercise 1
```{r ex1.1}
sex_survival <- donner %>% group_by(sex) %>% mutate(survived = 
ifelse(survived == 0, "Did not Survive", "Survived"))
```

```{r ex1.2}
donner %>% group_by(sex) %>%
  ggplot(data = sex_survival, mapping = aes(x = sex, fill = survived)) + 
  geom_bar(position = "fill") + labs(title = "Proportion of Population versus 
                                     Gender", x = "Gender", y = "Population 
                                     Proportion", fill = "Survived")
```

Since the proportion of the population that survived is greater for women 
than men, the relationship between sex and survival is that females have a 
higher chance of survival than males. 

## Exercise 2
```{r ex2}
age_survival <- donner %>%
  mutate(survived = ifelse(survived == 0, "Did not Survive", "Survived"))
age_survival

ggplot(data = age_survival, mapping = aes(x = survived, y = age)) + 
  geom_boxplot() + labs(title = "Age versus Survival", x = "Survived?", 
  y = "Age")
```
Since the median age of individuals that did not survive is greater than the 
median age of those who did survive, it can be inferred that younger individuals
are more likely to survive than older individuals. 

## Exercise 3
```{r ex3}
log_mod <- glm(survived ~ sex + age, data = donner, family = "binomial")
tidy(log_mod)
```

## Exercise 4
Logistic regression model: 
log(p/(1-p)) = 0.39237 + 1.20679sex - 0.033798age

## Exercise 5
In the context of the problem, this is the baseline or the y-intercept which 
means this indicates the odds of survival given that the individual is male and 
age 0. So, with age and sex held constant, the expected odds of survival is 
1.48049. 

## Exercise 6
In the context of the problem, this indicates how, holding sex constant, the 
odds of survival are for each year increase on average for males. It can be 
said that the odds of survival are expected to be multiplies by 0.9668 for 
each increase in age by 1 year. 

## Exercise 7
In the context of the problem, this indicates how the odds of survival compare 
for females over males when holding age constant. The odds of survival for 
females is 3.3427 times more likely than for males when holding age constant. 

## Exercise 8
For a 60 year old man:
```{r ex8.1}
tibble(
  age = 60,
  sex = factor("Man")) %>%
  augment(log_mod, newdata = .) %>%
  mutate(p = exp(.fitted) / (1 + exp(.fitted))) %>%
  pull(p)
```
Based on this logistic model, the predicted probability of survival for a 60 
year old man is 0.1630737.

For a 20 year old man:
```{r ex8.2}
tibble(
  age = 20,
  sex = factor("Man")) %>%
  augment(log_mod, newdata = .) %>%
  mutate(p = exp(.fitted) / (1 + exp(.fitted))) %>%
  pull(p)
```
Based on this logistic model, the predicted probability of survival for a 20 
year old man is 0.4295716.

For a newborn female:
```{r ex8.3}
tibble(
  age = 0,
  sex = factor("Woman")) %>%
  augment(log_mod, newdata = .) %>%
  mutate(p = exp(.fitted) / (1 + exp(.fitted))) %>%
  pull(p)
```
Based on this logistic model, the predicted probability of survival for a 
newborn (O year old) female is 0.8319002.


## Exercise 9
```{r}
tibble(
  age = rep(1:70, times = 2),
  sex = rep(c("Man", "Woman"), times = 70)) %>%
    augment(log_mod, newdata = .) %>%
    mutate(predicted_p = exp(.fitted) / (1 + exp(.fitted))) %>%
             ggplot(aes(x = age, y = predicted_p, color = sex)) + 
             geom_line(size = 1.5) + theme_minimal(base_size = 14) + 
             scale_color_manual(values = c("blue", "red")) + labs(title = 
            "Predicted Probability Values versus 
            Age Graph", subtitle = "Colored by Gender", x = "Age", y = 
            "Predicted Probability of Survival", color = "Gender")

```
Based on this graph, it is apparent that as age increases, the predicted 
probability of survival decreases regardless of gender. Women always have a 
higher probability of 
survival than men regardless of age, but since the curve of best fit for both 
males and females decrease at relatively the same rate, the probability decrease
with age increase is not selective for males or females. 

## Exercise 10
p = (e^(B0+B1x1+B2x2+...+Bkxk))/(1 + e^(B0+B1x1+B2x2+···+Bkxk)) > 0.75
p = (e^(0.39237 + 1.20679sex − 0.033798age))/(1 + e^(0.39237+1.20679 sex− 
0.033798age)) > 0.75
e^(0.39237 + 1.20679sex - 0.033798age) = 0.75 + 0.75e^(0.39237 + 1.20679sex - 0.033798age)

Since the gender of this member is female, the dummy variable sex is 1. 
e^(0.39237+1.20679(1)−0.033798 age) = 0.75 + 
0.75e^(0.39237+1.20679(1)−0.033798age)
e^(1.60006−0.033798age) = 0.75 + 0.75e^(1.60006−0.033798age)

After subtracting the common terms from both sides, and dividing each side by 
0.25:
0.25e^(1.60006−0.033798age) = 0.75
e^(1.60006−0.033798age) = 3

Then taking the natural log of each side and solving for age:
ln(e^(1.60006−0.033798 age)) = ln(3) 
1.60006 − 0.033798 age = 1.0986122 
1.60006 − 0.033798 age = 1.0986122 

−0.033798 age = −0.5014478 
age = 14.8366116 

Based on the logistic regression model, a female member of the Donner party has 
to be under 14.84 years to have a predicted probability of survival higher than 
0.75. But, since age is of type int, this is rounded down to age 14 so the drawn conclusion is that a female Donner party member must be less than 14 years old 
in order to have a predicted probability of survival greater than 0.75. 

## Exercise 11
Given the data, there are a few limitations. One, is that since there are 34 
females and 56 males, the sex variable has an unequal distribution and could 
cause a skew in the data. 
Also, there are linearity assumptions -- there is assumed independence between 
variables age, gender and survival. However, these variables could have some 
sort of relationship, such as say between gender and survival in that females scientifically have longer life spans than males. The coefficient estimates 
could therefore be unreliable if the predictors are correlated due to 
confounding variables (another being if the individual has pre-existing health 
conditions and therefore less chance of survival).
